---
title: "Stat 5309 Final Exam"
author: "Tom Wilson"
date: "May 7th, 2019"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(agricolae)
library(gmodels)
```


# Problem 1:
The effect of 4 types of graphite coater types on light box reading are to be studied. The readings might differ from day to day. Here assume we have a fixed effect model. Observations are taken for 3 days on the four types. The results are

```{r}
days<-c('d1','d2','d3')
types<-c('t1','t2','t3','t4')
graphite_data <- expand.grid(day=days,type=types)
graphite_data <- cbind(graphite_data,light=c(4,4.8,4,
                                             4.8,5,4.8,
                                             5,5.2,5.6,
                                             4.6,4.6,5))
graphite_data %>% kable()
```

## (a)
Note: Day is used as blocking factor

Write the model equation 
Let y be the response of light, b be Day (the blocking factor), and tau be the type treatment.
i and j are the level of day and type respectively. I and J are the number of levels, n is the number of observations. $\tau_i,\beta_j$ are treatment effects and block effects.

$$ y_{ij} = \mu + b_i + \tau_j + \epsilon_{ij} $$


State all the assumptions about residuals  

This model assumes that the error term is normally distributed with mean of zero and homogenous variance of $\sigma^2$ across levels of the treatment and block.

## (b)
Use definitions to estimate the model parameters ( 14 parameters)

The grand mean 
$$ \mu_{..} = \frac{1}{n} \sum_{i=1}^{4} \sum_{j=1}^{3}  y_{ij} = \frac{57.4}{12} =  4.78 $$



$$ \mu_1 = \frac{1}{I}\sum_{i=1}^3{y_{i,1}} = \frac{12.8}{3} = 4.267  $$
$$ \mu_2 = \frac{1}{I}\sum_{i=1}^3{y_{i,2}} = \frac{14.6}{3} = 4.867 $$

$$\mu_3 = \frac{1}{I}\sum_{i=1}^3{y_{i,3}} = \frac{15.8}{3} = 5.267  $$

$$\mu_4 \frac{1}{I}\sum_{i=1}^3{y_{i,4}} = \frac{14.2}{3} = 4.733   $$ 

$$\tau_1  = \mu_1 - \mu_{..}  = 4.267 - 4.78 = -0.513 $$
$$ \tau_2 = \mu_2 - \mu_{..}  = 4.867 - 4.78 = 0.087 $$

$$\tau_3 = \mu_3 - \mu_{..}  = 5.26 - 4.78 = 0.48 $$


$$\tau_4 = \mu_3 - \mu_{..}  = 4.733 - 4.78 =  -0.047 $$

$$\sigma^2 = \frac{1}{(I-1)(J-1)} \sum_{i=1}^{4} \sum_{j=1}^{3}  \epsilon_{ij}^2  = \frac{1}{(I-1)(J-1)} \sum_{i=1}^{4} \sum_{j=1}^{3}  (y_{ij}-\mu_{i})^2 = \frac{0.54}{6} = 0.09 $$

$$\beta_1 = \sum_{j=1}^{4} y_{1j} - \mu_{..}  = \frac{18.4}{4} - 4.78 = -0.18 $$

$$\beta_2 = \sum_{j=1}^{4} y_{1j} - \mu_{..}  = \frac{19.6}{4} - 4.78 = 0.12 $$

$$\beta_3 = \sum_{j=1}^{4} y_{1j} - \mu_{..}  = \frac{19.4}{4} - 4.78 = 0.07 $$


## (c)

Test the hypothesis that the effects ( ie, mean effect) of the 4 graphite coater types are the same. 

Hypothesis:  Ho: $$ \tau_1 = \tau_2 = \tau_3 = \tau_4 = 0 $$ Ha: one of $$ \tau_1,\tau_2,\tau_3,\tau_4 \neq 0 $$

Test statistic F=5.67

P-values = 0.03481

Conclusion: at a confidence level of 95% we would reject the joint null hypothesis and conclude that at least one effect is significantly different from zero. 

## (d)
How would your analysis of variance be different if the experiment had not been blocked?

If the experiement were not blocked, our best estimate of $\sigma^2$ would be a pooled variance which is likely to be much larger.

Write the ANOVA table with the value of the test statistic and p-value.

|         | degrees of freedom | sum of squares | mean square | F stat | P value |
|---------|--------------------|----------------|-------------|--------|---------|
| day     | 3-1=2              | 0.2067         | 0.1033      | 1.1481 | 0.37827 |
| type    | 4-1=3              | 1.53           | 0.51        | 5.667  | 0.035   |
| residual| 12-2-3-1=6         | 0.54           | 0.09        | 


# Problem 2:
The response time in milliseconds was determined for 3 different types of circuits that could be used in an automatic valve shutoff mechanism. The results from a completely randomized experiment are shown in the following table:

```{r}
circuit_types <- c('c1','c2','c3')
circuit_data <- data.frame(circuit_type=rep(circuit_types,5),
                           response_time=c(9,20,6,
                                           12,21,5,
                                           10,23,8,
                                           8,17,16,
                                           15,30,7))
circuit_data
```

## (a)
Write the model equation: $$ y = \mu + \tau_i + \epsilon_i $$

State all the assumptions about residuals

This model assumes that the error term is normally distributed with mean of zero and homogenous variance of $\sigma^2$ across levels of the treatment.
$\mu$ is the grand mean
$\tau_i$ are treatment effects



## (b)
```{r}
circuit_data %>% group_by(circuit_type) %>% summarise(n=length(response_time),s=sum(response_time),mu=mean(response_time))
```
Use definitions and R to find the estimates for the following parameters:
$$ \mu = \frac{1}{n} \sum_{i=1}^{3} \sum_{r=1}^{5} y_{ir} = \frac{207}{15} =  13.8 $$
$$ \mu_1 = \frac{1}{R} \sum_{r=1}^{5} y_{1,r} = \frac{54}{5} =  10.8 $$
$$ \mu_2 = \frac{1}{R} \sum_{r=1}^{5} y_{2,r} = \frac{111}{5} =  22.2 $$
$$ \mu_3  = \frac{1}{R} \sum_{r=1}^{5} y_{1,r} = \frac{42}{5} =  8.4 $$
$$ \tau_1 = \mu_1 - \mu = 10.8 - 13.8 = -3 $$

$$ \tau_2 = \mu_2 - \mu = 22.2 - 13.8 = 8.4 $$
$$ \tau_3 = \mu_3 - \mu = 8.4 - 13.8 = -5.4 $$

$$ \sigma^2 = \frac{1}{(n-I)} \sum_{i=1}^{3} \sum_{r=1}^{5}  \epsilon_{ir}^2  = \frac{1}{(n-I)} \sum_{i=1}^{3} \sum_{r=1}^{5}  (y_{ir}-\mu_{i})^2 = \frac{202.8}{12} = 16.9$$

```{r}
circuit_model <- aov(formula = response_time~circuit_type,data=circuit_data)
anova(circuit_model)
```

## (c)
Perform a Hypothesis Testing that the hypothesis to be tested for response time being equal for different circuit types.
Hypothesis.:  Ho: $$ \tau_1 = \tau_2 = \tau_3 = 0 $$ Ha: one of $$ \tau_1,\tau_2,\tau_3 \neq 0  $$

Test statistics F = 16.083

P-value = 0.0004

Decision: At a confidence of 95%, we reject the null hypothesis and conclude that at least one effect is not zero.

## (d)
Identify the pairs of treatment means which are different using Tukey’s tests with Significant level.

circuit types 2 and 1 are different 

circuit types 3 and 2 are different

circuit types 3 and 1 are similar

```{r}
TukeyHSD(circuit_model)
```

## (e)
Suppose we want to test for the significant difference of

$\frac{\mu_1 + \mu_3}{2} = \mu_2$

Perform the appropriate test for the contrast.

Write Contrast  C: 1,-2,1

Find the estimate :  c = -25.2 se(c) = 4.5
```{r}
fit.contrast(circuit_model,'circuit_type',coeff = c(1,-2,1))
```

Hypothesis :  Ho: $\mu_2$ is the average of $\mu_2$ and $\mu_3$; Ha:  $\mu_2$ <> average of $\mu_2$ and $\mu_3$

Test statistics -5.6 
P-value 0.00012

Conclusion: Reject the null hypothesis and conclude that $\mu_2$ is not the average of $\mu_1$ and $\mu_3$

# Problem 3
A study is conducted to compare 4 menus in terms of numbers of calories.
The 4 menus are:
 A- No calories
 B- Calories
 C-Rank-ordered Calories
 D-Color-ordered Calories

Suppose n=20  per each menu.  The sample means and estimated variance are

$y_1 = 17.6$ 
$y_2 = 16.8$
$y_3 = 16$
$y_4 = 14.4$
$y_..= 16.2$
$s^2 = MSE = 196 $

## (a)
Complete the following table to test the $H_0 : \mu_1 = \mu_2 = \mu_3 = \mu_4$

```{r}
df <- 79 - 3
mse <- 196
sserror <- mse * df
sstotal <- (17.6)^2*20 + (16.8)^2*20 + (16)^2*20 + (14.4)^2*20 - 16.2^2/80
sstrt <- sstotal - sserror
```

```{r}
1 - pf(10.55,3,76)
```

|source   | SS          | df     | MS    | F    | P-value |
|---------|-------------|--------|-------|------|---------|
| trt     | 6207.919    | 3      | 2069.3| 10.55| 7.05e-6 |
| error   | 14896       | 76     | 196   |      |         |
| total   | 21103.9     | 79     |       | 



Reject the null

## (b)
Difference between 2 means: $y_1 - y_2$

$$ SE = \sqrt{\frac{\sigma^2}{n}} = \sqrt{\frac{196}{80}} = 1.565 $$

Compute the standard error : 1.565

```{r}
t <- qt(1-0.025,40)
t
```

```{r}
2.02*1.565*0.3162
```

$$ LSD = t * SE * \sqrt{\frac{1}{20} + \frac{1}{20}} = 2.02*1.565*0.3162 = 0.9996 $$
Compute the Fisher LSD 0.9996


Write a 95%-CI for the difference, using Fisher LSD 

$ -0.9996 < y_1 - y_2 < 0.9996 $

## (c)
Contrast  C:
Write the contrast C which compares Menu 1 vs Menus {1,2,3} combined. 

$$\mu_1 = \frac{\mu_1 + \mu_2 + \mu_3 }{3}$$

$$3\mu_1 = \mu_1+ \mu_2 + \mu_3 $$

$$0 = \mu_1 -3\mu_1 + \mu_2 + \mu_3 $$

(-2,1,1,0)
                
Calculate an estimate of C :  c=-2.4
```{r}
-2*17.6+16.8+16.0
```
  

Calculate the standard error of the contrast: se(c ) =3.13

```{r}
sqrt(4*1.565^2)
```

Write a 95%-CI for  C, using t-distribution :(-6.26,6.26)
```{r}
qt(1-0.025,60)*3.13
```

# Problem 4

A router is used to cut locating notches on a printed circuit board. The vibration level at the surface of the board as it is cut is considered to be a major source of dimensional variation in the notches. Two factors are thought to influence vibration: bit size (A) and cutting speed (B). Two sizes (1/16 and 1/8 inch) two speeds (40 and 90 rpm ) are selected and four boards are cut at each set of conditions shown below. The response variable is vibration measuerd as the resultant vecor of three accelerometers (x,y,z) on each test circuit board.

```{r}
bit_sizes <- c(0,1)
speeds <- c(0,1)
vibration_data <- expand.grid(bit_size = bit_sizes,
                              speed = rep(speeds,4))
interactions =vibration_data$bit_size * vibration_data$speed
vibration_data <- cbind(vibration_data,
                        interaction = interactions,
                        vibation=c(18.2,27.2,15.9,41.0,
                                   18.9,24.0,14.5,43.9,
                                   12.9,22.4,15.1,36.3,
                                   14.4,22.5,14.2,39.9
                                   ))
vibration_data %>% kable()
```

## (a)
Calculate, from scratch, effects of A, B, AB[ Hint: Chapter 6]

```{r}
n <- 4
one <- (18.2+18.9+12.9+14.4)
a <- (27.2+24.0+22.4+22.5)
b <- (15.9+14.5+15.1+14.2)
ab <- (41.0+43.9+36.3+39.9)
A <- 1/(2*n) * ((ab-b)+(a-one))
B <- 1/(2*n) * ((ab-a)+(b-one))
AB <- 1/(2*n) * ((ab-b)+(a-one)) 
A
B
AB
```
A = 16.6375
B = 7.5375
AB = 16.6

## (b)
Calculate the Sum Squares: SS(A), SS(B), SS(AB).

```{r}
SSA  <- (ab+a-b-one)^2/(4*n)
SSB  <- (ab+b-a-one)^2/(4*n)
SSAB <- (ab+one-a-b)^2/(4*n)
SSA
SSB
SSAB
```
SSA = 1107.226
SSB = 227.2556
SSAB = 303.6306

## (c)
Complete the ANOVA table: df, MS, F-stat, P-values of A, B, AB
```{r}
SST <- sum(vibration_data$vibation^2) - mean(vibration_data$vibation)^2/(4*n)
SSE <- SST - SSA - SSB - SSAB
SST
SSE
```
SST = 10761.19
SSE = 9123.083

|source   | SS          | df    | MS     | F     | P-value |
|---------|-------------|-------|--------|-------|---------|
| A       | 1107.23     | 1     | 1107.23| 1.456 |<0.01    |
| B       | 227.26      | 1     | 227.26 | 0.298 |0.85     |
| AB      | 303.63      | 1     | 303.63 | 0.399 |0.711    |
| error   | 9123.083    | 12    | 760.26 |       |         |
| total   | 10761.19    | 15    |        |       |         |



## (d)
Use R,  find the model matrix X.

```{R}
model_matrix <- cbind(1,vibration_data[,c('bit_size','speed','interaction')])
X <- model_matrix %>% as.matrix
y <- vibration_data$vibation
X %>% kable()

```

Use X to calculate the vector of  coefficients   ,  for the regression model.

```{r}
solve(t(X)%*%X) %*% t(X) %*% y
```
$B_0 = 16.1$
$B_A = 7.925$
$B_B = -1.175$
$B_AB = 17.425$