---
title: "Stat 5309 Final Exam"
author: "Tom Wilson"
date: "May 7th, 2019"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(agricolae)
```


# Problem 1:
The effect of 4 types of graphite coater types on light box reading are to be studied. The readings might differ from day to day. Here assume we have a fixed effect model. Observations are taken for 3 days on the four types. The results are

```{r}
days<-c('d1','d2','d3')
types<-c('t1','t2','t3','t4')
graphite_data <- expand.grid(day=days,type=types)
graphite_data <- cbind(graphite_data,light=c(4,4.8,4,
                                             4.8,5,4.8,
                                             5,5.2,5.6,
                                             4.6,4.6,5))
graphite_data %>% kable()
graphite_data %>% write.csv("exam01graphite_data.csv")
```

## (a)
Note: Day is used as blocking factor

Write the model equation 

State all the assumptions about residuals  


## (b)
Use definitions to estimate the model parameters ( 14 parameters)
$$ 
\mu_1, 
\mu_2, 
\mu_3, 
\mu_4, 
\mu_5,
\tau_1, 
\tau_2,
\tau_3,
\tau_4,
\sigma^2,
\beta_1,
\beta_2,
\beta_3,
\beta_4
$$
where $\tau_i,\beta_j$ are treatment effects and block effects.

## (c)

Test the hypothesis that the effects ( ie, mean effect) of the 4 graphite coater types are the same. 

Hypothesis:  Ho________________Ha__________

Test statistic______________

P-values _________________

Conclusion_________________

## (d)
How would your analysis of variance be different if the experiment had not been blocked?

Write the ANOVA table with the value of the test statistic and p-value.


# Problem 2:
The response time in milliseconds was determined for 3 different types of circuits that could be used in an automatic valve shutoff mechanism. The results from a completely randomized experiment are shown in the following table:

```{r}
circuit_types <- c('c1','c2','c3')
circuit_data <- data.frame(circuit_type=rep(circuit_types,5),
                           response_time=c(9,20,6,
                                           12,21,5,
                                           10,23,8,
                                           8,17,16,
                                           15,30,7))
circuit_data
circuit_data %>% write.csv("exam02circuit_data.csv")
```

## (a)
Write the model equation:_________________________________
State all the assumptions about residuals:  _________________________

## (b)
Use definitions and R to find the estimates for the following parameters:
$$
\mu_1,
\mu_2,
\mu_3,
\mu,
\tau_1,
\tau_2,
\tau_3,
\sigma^2
$$
$\tau_i$ are treatment effects

## (c)
Perform a Hypothesis Testing that the hypothesis to be tested for response time being equal for different circuit types.
Hypothesis.:  Ho__________________ Ha_______________

Test statistics _______________

P-value____________________

Decision: __________________ Use 

## (d)
Identify the pairs of treatment means which are different using Tukey’s tests with Significant level.


## (e)
Suppose we want to test for the significant difference of

$\frac{\mu_1 + \mu_3}{2} = \mu_2$

Perform the appropriate test for the contrast.

Write Contrast  C: ____________________

Find the estimate :  c =_____________se(c ) =__________________

Hypothesis :  Ho__________________; Ha________________

Test statistics_____________________; P-value______________

Conclusion:______________________

# Problem 3
A study is conducted to compare 4 menus in terms of numbers of calories.
The 4 menus are:
 A- No calories
 B- Calories
 C-Rank-ordered Calories
 D-Color-ordered Calories

Suppose n=20  per each menu.  The sample means and estimated variance are

$y_1 = 17.6$ $y_2=16.8$ $y_3=16$ $y_4=14.4$ $y_..=16.2$ $s^2 = MSE = 196 $

## (a)
Complete the following table to test the $H_0 : \mu_1 = \mu_2 = \mu_3 = \mu_4$

Reject ____________ Not reject ______________

## (b)
Difference between 2 means: $y_1 - y_2$

Compute the standard error :_____________

Compute the Fisher LSD_____________


Write a 95%-CI for the difference, using Fisher LSD____________________

## (c)
Contrast  C:
Write the contrast C which compares Menu 1 vs Menus {1,2,3} combined. ___________________________________
                
Calculate an estimate of C :  c=_________________
  

Calculate the standard error of the contrast: se(c ) =______________


Write a 95%-CI for  C, using t-distribution :___________________

# Problem 4

A router is useed to cut locating notches on a printed circuit board. The vibration level at the surface of the board as it is cut is considered to be a major source of dimensional variation in the notches. Two factors are thought to influence vibration: bit size (A) and cutting speed (B). Two sizes (1/16 and 1/8 inch) two speeds (40 and 90 rpm ) are selected and four boards are cut at each set of conditions shown below. The response variable is vibration measuerd as the resultant vecor of three accelerometers (x,y,z) on each test circuit board.

```{r}
bit_sizes <- c('1/16th','1/8th')
speeds <- c('40','90')
vibration_data <- expand.grid(bit_size = bit_sizes,
                              speed = rep(speeds,4))
vibration_data <- cbind(vibration_data,
                        vibation=c(18.2,27.2,15.9,41.0,
                                   18.9,24.0,14.5,43.9,
                                   12.9,22.4,15.1,36.3,
                                   14.4,22.5,14.2,39.9
                                   ))
vibration_data %>% kable()
vibration_data %>% write.csv("exam04vibration_data.csv")
```

## (a)
Calculate, from scratch, effects of A, B, AB[ Hint: Chapter 6]

## (b)
Calculate the Sum Squares: SS(A), SS(B), SS(AB).

## (c)
Complete the ANOVA table: df, MS, F-stat, P-values of A, B, AB

## (d)
Use R,  find the model matrix X.

Use X to calculate the vector of  coefficients   ,  for the regression model.

