---
title: "Stat 5309 Lab 1"
author: "Tom Wilson"
date: "January 18, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.
Data: Bottle Filling Machines. 
Two machines are used to fill plastic bottles with a volume of 16.0 onces.
The process can be assumed to be normal, with $\sigma_1 = 0.015$, $\sigma_2 =0.018$. An experiment is performed.

```{r}
machine_1 <- c(16.03, 16.01, 16.04, 15.96, 16.05,
               15.98, 16.05, 16.02, 16.02, 15.99
               )

machine_2 <- c(16.02, 16.03, 15.97, 16.04, 15.96,
               16.02, 16.01, 16.01, 15.99, 16.00
               )
```

## a
Test hypothesis: $H_0: \mu_1 = \mu_2$ . 

use a z-test because the process is normal and the variance(s) are known.

```{r}
n <- length(machine_1)
machine_1_sigma <- 0.015
machine_2_sigma <- 0.018
pooled_sigma <- sqrt(machine_1_sigma^2 + machine_2_sigma^2)
SE = pooled_sigma/sqrt(n)

machine_1_mu <- mean(machine_1)
machine_2_mu <- mean(machine_2)
mu_diff <- machine_1_mu - machine_2_mu

CL <- 0.95
alpha = 1 - CL
half_alpha = alpha/2
z = mu_diff/SE
p_value = 1 - pnorm(z)
mu_diff
p_value
```
The difference between the means is 0.01.
Given the null hypothesis, the probability of observing a difference at least this extreme is 8.8%. at a confidence level of 5% we would fail to reject the null hypothesis and conclude that there is no difference between machine 1 and machine 2.


## b
Find the P-value for the test.

The p-value is calculated above as 8.8%

## c
Find the 95%-CI for the difference of the population means.

```{R}
z_critical <- qnorm(1-half_alpha)
CI=c(mu_diff-z_critical*SE , mu_diff+z_critical*SE)
CI
```
the difference between the means is between -0.0045 and 0.0245.
Since this interval contains 0, it is consistent with the conclusion in part a.

## d
Perform a boxplot and a stripchart with line connecting means.


```{R}
y <- data.frame(machine_1,machine_2)
names(y) <-c("machine_1", "machine_2")
y.means <- apply(y,2,mean)
stripchart(y, xlab="oz", vertical=TRUE,pch=16)
lines(y.means)
```
```{R}
boxplot(y)
```

# 2.
Data: Burning times of Chemical Flares.

```{r}
type_1 <- c(65,82,81,67,57,59,66,75,82,70)

type_2 <- c(64,56,71,69,83,74,59,82,65,79)
```

**Perform by R . Only use var.test() to check**

## a
Test the hypothesis that the two variances are equal. $\alpha = 0.05$

```{R}
var.test(type_1,type_2)
```

under the null hypothesis of equal variance, we would expect would expect an F stat this extreme 97% of the time. At a confidence of 5%, we fail to reject the null hypothesis and conclude that the variance of test1 is the same as that of test2. 

## b
Test the hypothesis that the burning times are equal. 

```{R}
t.test(type_1,type_2,var.equal = TRUE)
```

Test statistics, t = 0.048 with 18 degrees of freedom

P-value = 0.9622

Conclusion.

Given that the means are equal, we would expect a difference at least this great 96% of the time. At a confidence level of 5% we would fail to reject the null hypothesis and conclude that the means are indeed the same.

## c
Check the normality assumption for both types of flares.

```{R}
qqnorm(type_1,main="Quantile-Quantile plot for Type 1")
qqline(type_1)
```

```{r}
qqnorm(type_2,main="Quantile-Quantile plot for Type 1")
qqline(type_2)
```

Visually, the quantiles of both type1 and type2 closely match that of a normal distribution.