---
title: "Stat 5309 Lab 1"
author: "Tom Wilson"
date: "January 18, 2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(agricolae)
```

# 1  Data: Bacteria with Packages

## a 
Set up th dataframe

```{r}
bacteria <- data.frame(package = rep(c("plastic","vacuum","1%CO","100%CO2"),each=3),
                       log_count = c(7.66,6.98,7.80,
                                     5.26,5.44,5.80,
                                     7.41,7.33,7.04,
                                     3.51,2.91,3.66))


bacteria %>% kable()
```

## b
perform a stripchart with line connecting means of logcount vs package

```{r}
stripchart(bacteria$log_count~bacteria$package,vertical = TRUE)

lines(tapply(bacteria$log_count,bacteria$package , mean))


```

## c
build a linear model using aov() response as logcount. Do a summary.lm() and summary.aov()

```{r}
fit <- lm(data=bacteria,formula = log_count ~ package)
summary_of_fit <- summary(fit)
anova_of_fit <- anova(fit)
anova_of_fit
aov(fit)
```

## d

perform a bartlett test of equal variances

```{r}
bartlett.test(bacteria$log_count~bacteria$package)
```

## e
perform a multiple comparison of treatment mean using TukeyHSD()

```{r}
TukeyHSD(aov(fit), conf.level=0.95)
```

# 2 Data: Tensile strength of Portland Cement
## a 
Set up a data frame with variables mixing and strength

```{r}
cement <- data.frame(  mixing = rep(c("1","2","3","4"),each=4),
                     strength = c(3129,3000,2865,2890,
                                  3200,3300,2975,3150,
                                  2800,2900,2985,3050,
                                  2600,2700,2600,2765))

cement %>% kable()
```

## b
Perform a stripchart and boxplot
```{r}
stripchart(cement$strength~cement$mixing,vertical = TRUE)
lines(tapply(cement$strength,cement$mixing , mean))

```

```{r}
boxplot(cement$strength~cement$mixing)
```

## c
use Fisher Least Significant Difference (LSD) with $\alpha=0.05$ to make a comparison

```{r}
model <- aov(strength~mixing, data=cement)
model_anova <- anova(model)
MSError <- model_anova$`Mean Sq`[2]
LSDout <- LSD.test(model,"mixing", MSerror = MSError, console = TRUE)
plot(LSDout)
```

## d
Test the hypothesis that mixing techniques affect the strength of the cement. Use $\alpha=0.05$. What test do you use? Conclusion?

To test the hypothesis that there are no differences among the means, an F-test is appropriate.

```{r}
model_summary <- summary(model)
model_summary
```
```{r}
p_value <- model_anova$`Pr(>F)`[1]
p_value
```
The observed p-value of 0.0005 is much less than the rejection criteria of 0.05; therefore, we reject the null hypothesis that there are no differences among the means and conclude that at least one of the treatment groups has a mean that is significantly different from the others.

# 3
A manufacture of television sets is interested in the effect on tube conducivity of four different types of ocutaing for color picture tubes. the following conductivity data are obtained.

```{r}
television <- data.frame(coating_type = rep(c("1","2","3","4"),each=4),
                         conductivity = c(143,141,150,146,
                                          152,149,137,143,
                                          134,136,132,127,
                                          129,127,132,129))

television %>% kable()
```

## a 
Is there a difference in conductivity due to coating type? Use $\alpha=0.05$

```{r}

```

## b
Estimate the overal mean and the treatment effects.

```
```

## c
Compute a 95% confidence interval estimate of th mean of coating type 4. 
```
```

Compute a 99% confidence interval estimate of the mean difference between coating types 1 and 4.
```
```

## d
Test all pairs of means using the Fisher LSD method with $\alpha=0.05$
```
```

## e
Use the graphical method discussed in Section 3-5.3 to compare the means. Which coating type produces the highest conductivity?
```
```

## f
Assuming that coating type  is currently in use, what are your recommendations to the manufacturer? We wish to minimize conductivity.
