---
title: "Stat 5309 Lab 4b"
author: "Tom Wilson"
date: "Feb 21, 2019"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(agricolae)
```

# 1.

The yield of a chemical process is being studied. The two most important variables are thought ot be th pressure and the temperature. Three levels of each factor are selected and a factorial experiment with two replicates is performed. The yield data follows.

## a
Set up the dataframe.

```{r}
temperatures <- c("150","160","170")
pressures <- c("200","215","230")
process <- expand.grid(pressure = rep(pressures,2),
                       temperature = temperatures)

process <- cbind(process,yield=c(90.4,90.7,90.2,
                                 90.2,90.6,90.4,
                                 90.1,90.5,89.9,
                                 90.3,90.6,90.1,
                                 90.5,90.8,90.4,
                                 90.7,90.9,90.1))
process %>% kable()
```

## b
Build a linear model using aov(). Are the pressure means significant? Are the temp means significant? Is the interaction significant?

## c
Create a boxplot of 
yield vs temp
yield vs pressure
yield vs temp and pressure

## d
Perform a residuals assumption check

# 2
Johnson an dLeone describe an experiment to investigate warping of copper plates. The two factors studied were the temperature and the copper content of the plates. The response variable was a meaure of the amont of warping. The data were as follows.

## a
Set up the dataframe 

```{r}
temperatures <- c("50","75","100","125")
copper_contents <- c("40","60","80","100")
copperplate <- expand.grid(copper_content=rep(copper_contents,2),
                           temperature=temperatures
                           )
copperplate <- cbind(copperplate,warping = c(17,16,24,28,
                                             20,21,22,27,
                                             12,18,17,12,
                                             9,13,12,31,
                                             16,18,25,30,
                                             12,21,23,23,
                                             21,23,23,29,
                                             17,21,22,31))
copperplate
```

## b
build a response model surface (RSM) with warpage as response, use rsm().


