---
title: "Stat 5309 Midterm Project"
author: "Tom Wilson"
date: "Mar 21, 2019"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(agricolae)
library(rsm)
```


# 1
The effective life of insulating luids at an accelerated load of 35 kV is being studied.
Test fluid_data have been obtained for four types of fluids. The results were as follows:

## a
Either read fluid_data into R or create the fluid_dataframe.

```{r}
fluidtypes <- c("1","2","3","4")
fluid_data <- data.frame(fluid = rep(fluidtypes,each=6)
                   ,lifetime=c(17.6,18.9,16.3,17.4,20.1,21.6,
                               16.9,15.3,18.6,17.1,19.5,20.3,
                               21.4,23.6,19.4,18.5,20.5,22.3,
                               19.3,21.1,16.9,17.5,18.3,19.8
                               ))
fluid_data %>% kable()
```

## b
Build a linear model, using aov. Is there a significant difference among treatment means?
which fluid gives the longer life?

```{r}
insulation_life_model <- aov(formula = lifetime ~ fluid,data=fluid_data)
summary(insulation_life_model)
```
Given that there is no difference in lifetime between fluid types, we would expect a result at least this extreme 5.25% of the time. At a confidence level of 95% we fail to reject the null hypothesis and conclude that any difference observed is due to chance.

## c
Construct a 95% Confidence Interval for themean life of fluid 2.

```{r}
anova_of_insulation_life <- anova(insulation_life_model)
MSError=anova_of_insulation_life$`Mean Sq`[2]
LSD.test(insulation_life_model, "fluid", MSerror = MSError,console = TRUE)
```

16.40 to 19.50 is a 95% confidence interval for the mean of fluid type 2.


Construct a 99% Confidence Interval for the difference between the lives of Fluids 2 and 3.

```{r}
TukeyHSD(insulation_life_model, conf.level=0.99)
```
The difference between type 4 and type 1 is between -0.7213 and 6.7213 with 99% confidence.

## d
Perform a complete 3-part residuals check.
```{r}
plot(insulation_life_model)
```

### Independence
Based on a plot of standardized residual vs factor level (in this case, fluid type) the residuals are independent of fluid type. 

### Normality
Based on a quantile-quantile plot te residuals are very close to normally distrbuted with a mean of zero.

### homoscedasticity
based on a plot of residual vs fitted values, variation in residual is constant.

## e
Calculate the number of replicates for a power of 0.99

```{r}
fluid_means <- fluid_data %>% group_by(fluid) %>% summarise(trt_mean = mean(lifetime))
power.anova.test(groups = 4, 
                 between.var = var(fluid_means$trt_mean),
                 within.var = anova_of_insulation_life$`Mean Sq`[2],
                 power=0.99
                 )
```
17 replicates are needed to achieve a power of 0.99 .

# 2

## a
Either read data into R or create the dataframe.
```{r}
oils <- c("1","2","3")
trucks <- c("1","2","3","4","5")
fuel_data <- expand.grid(truck=trucks,oil=oils)
fuel_data <- cbind(fuel_data,fuel_consumption = c(0.5,
                                        0.634,
                                        0.487,
                                        0.329,
                                        0.512,
                                        0.535,
                                        0.675,
                                        0.52,
                                        0.435,
                                        0.54,
                                        0.513,
                                        0.595,
                                        0.488,
                                        0.4,
                                        0.51))
fuel_data
```

## b
Build a linear model. Is there any significant difference of means about the oil types?
Which oil type gies the lowest fuel consumption?

```{r}
fuel_consumption_model <- aov(formula = fuel_data$fuel_consumption ~ oil+truck,data=fuel_data)
summary(fuel_consumption_model)
```

## c
Is the blocking approach effective?

## d
Do a complete residual assumption check.

# 3
Suppose that in in Problem 4-15, the engineer suspects that the workplaces used by the four operators may represent an additional source of variation. Analyze the data from this experiment (use $\alpha=0.05$) and draw conclusions.

## a
Set up a dataframe with 2 blocking factors (order and operator) and treatment (A,B,C,D)
```{r}
orders <- c("1st","2nd","3rd","4th")
operators <- c("op1","op2","op3","op4")
workplace_data <- expand.grid(operator=operators,order_of_assembly=orders)
workplace_data <- cbind(workplace_data,workplace=c("C","B","D","A",
                               "B","C","A","D",
                               "A","D","B","C",
                               "D","A","C","B"
                               ),
              observation=c(11,10,14,8,
                            8,12,10,12,
                            9,11,7,15,
                            9,8,18,6
                            )
              )
workplace_data %>% kable()
```

## b
Use Latin Square to analyze the treatment means.

## c
Which level combination brings the lowest time?

# 4
The factors that influence the breaking strength of a synthetic fiber are being studied. Four production machines and three operators are chosen and a factorial experiment is run using fiber from the same production batch. The results follow.

## a
Either read data into R or create the dataframe.
```{r}
machines <- c("1","2","3","4")
operators <- c("o1","o2","o3")
fiber_data <- expand.grid(machine=rep(machines,2),operator=operators)
fiber_data <- cbind(fiber_data,strength = c(109,110,108,110,
                                110,115,109,108,
                                110,110,111,114,
                                112,111,109,112,
                                116,112,114,120,
                                114,115,119,117
                                )
              )
fiber_data %>% kable()
```

## b
Build a linear model. Any interaction between operator and machine?

```{r}
fiber_model <- aov(formula = strength ~ operator+machine ,data=fiber_data)
summary(fiber_model)
```

## c
Build a reduced model.

```{r}
fiber_model_reduced <- aov(formula = strength ~ operator ,data=fiber_data)
summary(fiber_model_reduced)
```

## d
Do a complete 3-part residual assumption check.

# 5
An experiment is conducted to study the influence of operating temperature and three types of face-plate glass in the light output of an oscilloscope tube. The following data are collected.

## a
Either read data into R or create the dataframe.
```{r}
temperatures <- c(100,125,150)
glasses <- c("t1","t2","t3")
glass_data <- expand.grid(temperature=rep(temperatures,3),
                    glass = glasses)
glass_data <- cbind(glass_data,output=c(580,1090,1392,
                            568,1087,1380,
                            570,1085,1386,
                            
                            550,1070,1328,
                            530,1035,1312,
                            579,1000,1299,
                            
                            546,1045,867,
                            575,1053,904,
                            599,1066,889))
glass_data %>% kable()
```

## b
Build a linear model. Any interaction between glass type and temperature?

```{r}
light_output_model <- aov(formula = output~temperature+glass,data=glass_data)
summary(light_output_model)
```


## c
Build a reduced model.
```{r}
light_output_model <- aov(formula = output~temperature,data=glass_data)
summary(light_output_model)
```

## d
Do a complete 3-part residual assumption check.

# 6
Sludge is the dried product remaining from processed sewage. 
It can be used as fertilizer on crops. 
However, it contains heavy metals. 
They hypothesized the concentration of certain heavy metals in sludge differ among the metropolitan areas from which the sludge is obtained. 
The sludge was aded to the sand at 3 different rates: 0.5,1.0,1.5 metric tons/acre.
The zinc levels were recorded.

## a
Set up a dataframe named metals. Use factos city (A,B,C), rate (0.5,1.0,1.5), and zinc for the observations.
```{r}
cities <- c("A","B","C")
rates <- c(0.5,1.0,1.5)
zinc_data <- expand.grid(rate=rates,city=cities)
zinc_data <- cbind(zinc_data,zinc=c(26.4,25.2,26.0, 30.1,47.7,73.8, 19.4,23.2,18.9,
                          23.5,39.2,44.6, 31.0,39.1,71.1, 19.3,21.3,19.8,
                          25.4,25.5,35.5, 30.8,55.3,68.4, 18.7,23.2,19.6,
                          22.9,31.9,38.6, 32.8,50.7,77.1, 19.0,19.9,21.9
                          )
              )
zinc_data %>% kable()
```

## b
Build an aov model, using zinc as the response. Which factors are significant? Interaction is significant? Perform an interaction plot.

```{r}
zinc_model <- aov(formula = zinc ~ rate+city ,data=zinc_data)
summary(zinc_model)
```

## c
List all the factor means and effects. using tapply() or model.table().

## d
calculate the interaction sum squares from scratch.



<!-- 
 lab 5 below
-->

# 1.

The yield of a chemical process is being studied. The two most important variables are thought ot be th pressure and the temperature. Three levels of each factor are selected and a factorial experiment with two replicates is performed. The yield data follows.

## a
Set up the dataframe.

```{r}
temperatures <- c("150","160","170")
pressures <- c("200","215","230")
process <- expand.grid(pressure = rep(pressures,2),
                       temperature = temperatures)

process <- cbind(process,yield=c(90.4,90.7,90.2,
                                 90.2,90.6,90.4,
                                 90.1,90.5,89.9,
                                 90.3,90.6,90.1,
                                 90.5,90.8,90.4,
                                 90.7,90.9,90.1))
process %>% kable()
```

## b
Build a linear model using aov(). Are the pressure means significant? Are the temp means significant? Is the interaction significant?

```{r}
yield_model <- aov(yield ~ temperature * pressure, data=process)
summary(yield_model)
```
The interaction between temperature and pressure is not significant.
The main effects of temperature and pressure are significant.

```{r}
yield_model <- aov(yield ~ temperature + pressure, data=process)
summary(yield_model)
```

## c
Create a boxplot of

### yield vs temp
```{r}
boxplot(yield~temperature, data=process)
```

### yield vs pressure
```{r}
boxplot(yield~pressure, data=process)
```

### yield vs temp and pressure
```{r}
boxplot(yield~temperature*pressure, data=process)
```

## d
Perform a residuals assumption check
```{r}
qqnorm(yield_model$residuals)
qqline(yield_model$residuals)
```
The residuals are consistent with a normal distribution centered on zero with constant variance.

# 2
Johnson an dLeone describe an experiment to investigate warping of copper plates. The two factors studied were the temperature and the copper content of the plates. The response variable was a meaure of the amont of warping. The data were as follows.

## a
Set up the dataframe 

```{r}
temperatures <- c("50","75","100","125")
copper_contents <- c("40","60","80","100")
copperplate <- expand.grid(copper_content=rep(copper_contents,2),
                           temperature=temperatures
                           )
copperplate <- cbind(copperplate,warping = c(17,16,24,28,
                                             20,21,22,27,
                                             12,18,17,12,
                                             9,13,12,31,
                                             16,18,25,30,
                                             12,21,23,23,
                                             21,23,23,29,
                                             17,21,22,31))
copperplate %>% kable()
```

## b
build a response model surface (RSM) with warpage as response, use rsm().
```{r}
temp_nums <- copperplate$temperature %>% as.numeric()
cu_nums <- copperplate$copper_content %>% as.numeric()
response <- copperplate$warping
copper_model<- rsm(response ~ SO(temp_nums, cu_nums), data=copperplate)
summary(copper_model)

```
