---
title: "Stat 5309 Midterm Project"
author: "Tom Wilson"
date: "Mar 21, 2019"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(agricolae)
library(rsm)
```


# 1
The effective life of insulating fluids at an accelerated load of 35 kV is being studied.
Test fluid_data have been obtained for four types of fluids. The results were as follows:

## a
Either read data into R or create the dataframe.

```{r}
fluidtypes <- c("1","2","3","4")
fluid_data <- data.frame(fluid = rep(fluidtypes,each=6)
                         ,lifetime=c(17.6,18.9,16.3,17.4,20.1,21.6,
                                     16.9,15.3,18.6,17.1,19.5,20.3,
                                     21.4,23.6,19.4,18.5,20.5,22.3,
                                     19.3,21.1,16.9,17.5,18.3,19.8
                                     ))
fluid_data %>% kable()
```

## b
Build a linear model, using aov.

```{r}
insulation_life_model <- aov(formula = lifetime ~ fluid,data=fluid_data)
summary(insulation_life_model)
```
Is there a significant difference among treatment means?

Given that there is no difference in lifetime between fluid types, we would expect a result at least this extreme 5.25% of the time. At a confidence level of 95% we fail to reject the null hypothesis and conclude that any difference observed is due to chance.

which fluid gives the longer life?
```{r}
tapply(fluid_data$lifetime, fluid_data$fluid, mean)
```

Although it is not significant, fluid 3 is associated with the longest life.

## c
Construct a 95% Confidence Interval for themean life of fluid 2.

```{r}
anova_of_insulation_life <- anova(insulation_life_model)
MSError=anova_of_insulation_life$`Mean Sq`[2]
LSD.test(insulation_life_model, "fluid", MSerror = MSError,console = TRUE)
```

16.4 to 19.5 is a 95% confidence interval for the mean of fluid type 2.

Construct a 99% Confidence Interval for the difference between the lives of Fluids 2 and 3.

```{r}
TukeyHSD(insulation_life_model, conf.level=0.99)
```
The difference between type 4 and type 1 is between -0.7213 and 6.7213 with 99% confidence.

## d
Perform a complete 3-part residuals check.
```{r}
par( mfrow = c(2,2) )
plot(insulation_life_model)
```

### Independence
Based on a plot of standardized residual vs factor level (in this case, fluid type) the residuals are independent of fluid type. 

### Normality
Based on a quantile-quantile plot te residuals are very close to normally distrbuted with a mean of zero.

### Homoscedasticity
based on a plot of residual vs fitted values, variation in residual is constant.

## e
Calculate the number of replicates for a power of 0.99

```{r}
fluid_means <- fluid_data %>% 
               group_by(fluid) %>% 
               summarise(trt_mean = mean(lifetime))

power.anova.test(groups = 4, 
                 between.var = var(fluid_means$trt_mean),
                 within.var = anova_of_insulation_life$`Mean Sq`[2],
                 power=0.99
                 )
```
17 replicates are needed to achieve a power of 0.99 .

# 2

## a
Either read data into R or create the dataframe.
```{r}
oils <- c("1","2","3")
trucks <- c("1","2","3","4","5")
fuel_data <- expand.grid(truck=trucks,oil=oils)
fuel_data <- cbind(fuel_data,fuel_consumption = c(0.500,0.634,0.487,
                                                  0.329,0.512,0.535,
                                                  0.675,0.520,0.435,
                                                  0.540,0.513,0.595,
                                                  0.488,0.400,0.510
                                                  )
                   )
fuel_data
```

## b
Build a linear model. 

```{r}
fuel_consumption_model <- aov(formula = fuel_data$fuel_consumption ~ oil+truck,data=fuel_data)
summary(fuel_consumption_model)
```

Is there any significant difference of means about the oil types?

Both truck and oil have significantly different means.

Which oil type gies the lowest fuel consumption?

```{r}
tapply(fuel_data$fuel_consumption,fuel_data$oil,mean)
```

Oil type 1 is associated with the lowest fuel consumption.

## c
Is the blocking approach effective?

The blocking variable of truck type has a significant effect; therefore, blocking is effective. Without blocking, the effect of oil on fuel consumption could not be decoupled from trucktype.

## d
Do a complete residual assumption check.

```{r}
par( mfrow = c(2,2) )
plot(fuel_consumption_model)
```

### Independence
Based on a plot of standardized residual vs factor level (in this case, oil type) the residuals are nearly independent of oil type. 

### Normality
Based on a quantile-quantile plot te residuals are very close to normally distrbuted with a mean of zero.

### Homoscedasticity
based on a plot of residual vs fitted values, variation in residual is not constant. There may be a more complicated relationship between oil, truck, and fuel consumption.


# 3
Suppose that in in Problem 4-15, the engineer suspects that the workplaces used by the four operators may represent an additional source of variation. Analyze the data from this experiment (use $\alpha=0.05$) and draw conclusions.

## a
Set up a dataframe with 2 blocking factors (order and operator) and treatment (A,B,C,D)
```{r}
orders_3 <- c("1st","2nd","3rd","4th")
operators_3 <- c("op1","op2","op3","op4")
workplaces <- c("A","B","C","D")
workplace_data <- expand.grid(operator=operators_3,
                              order_of_assembly=orders_3
                              )
workplace_data <- cbind(workplace_data,
                        workplace=c("C","B","D","A",
                                    "B","C","A","D",
                                    "A","D","B","C",
                                    "D","A","C","B"
                                    ),
                        observation=c(11,10,14,8,
                                      8,12,10,12,
                                      9,11,7,15,
                                      9,8,18,6
                                      )
                        )
workplace_data %>% kable()
```

## b
Use Latin Square to analyze the treatment means.
```{r}
workplace_model <- aov(formula = observation ~ order_of_assembly+operator+workplace ,data=workplace_data)
summary(workplace_model)
```

At a confidence level of 95%, there are differences between workplaces.

```{r}
TukeyHSD( workplace_model, "workplace")
```

Pairwise, C is different from both B and A. Other pairs of workplaces are not significantly different from each other.

## c
Which level combination brings the lowest time?

```{r}
workplace_model_tables <- model.tables( workplace_model, type = "means" )
workplace_model_tables$tables$workplace
```
Workplace B is associated with the lowest time.

# 4
The factors that influence the breaking strength of a synthetic fiber are being studied. Four production machines and three operators are chosen and a factorial experiment is run using fiber from the same production batch. The results follow.

## a
Either read data into R or create the dataframe.
```{r}
machines <- c("1","2","3","4")
operators_4 <- c("o1","o2","o3")
fiber_data <- expand.grid(machine=rep(machines,2),operator=operators_4)
fiber_data <- cbind(fiber_data,strength = c(109,110,108,110,
                                            110,115,109,108,
                                            110,110,111,114,
                                            112,111,109,112,
                                            116,112,114,120,
                                            114,115,119,117
                                            )
                    )
fiber_data %>% kable()
```

## b
Build a linear model. Any interaction between operator and machine?

```{r}
fiber_model <- aov(formula = strength ~ operator*machine ,data=fiber_data)
summary(fiber_model)
```

There is no significant interaction between operator and machine. Only operator is significant.

## c
Build a reduced model.

```{r}
fiber_model_reduced <- aov(formula = strength ~ operator ,data=fiber_data)
summary(fiber_model_reduced)
```

## d
Do a complete 3-part residual assumption check.

```{r}
par( mfrow = c(2,2) )
plot(fiber_model_reduced)
```

### Independence
Based on a plot of standardized residual vs factor level (in this case, fluid type) the residuals are independent of fluid type. 

### Normality
Based on a quantile-quantile plot te residuals are very close to normally distrbuted with a mean of zero.

### Homoscedasticity
based on a plot of residual vs fitted values, variation in residual is constant.


# 5
An experiment is conducted to study the influence of operating temperature and three types of face-plate glass in the light output of an oscilloscope tube. The following data are collected.

## a
Either read data into R or create the dataframe.
```{r}
temperatures <- c(100,125,150)
glasses <- c("t1","t2","t3")
glass_data <- expand.grid(temperature=rep(temperatures,3),
                          glass = glasses)
glass_data <- cbind(glass_data,output=c(580,1090,1392,
                                        568,1087,1380,
                                        570,1085,1386,
                                        550,1070,1328,
                                        530,1035,1312,
                                        579,1000,1299,
                                        546,1045,867,
                                        575,1053,904,
                                        599,1066,889
                                        )
                    )
glass_data %>% kable()
```

## b
Build a linear model. Any interaction between glass type and temperature?

```{r}
light_output_model <- aov(formula = output~temperature*glass,data=glass_data)
summary(light_output_model)
```

There is a significant interaction between temperature and glass type.

## c
Build a reduced model.

Since the interaction is significant, the full interactive model is appropriate.

## d
Do a complete 3-part residual assumption check.

```{r}
par( mfrow = c(2,2) )
plot(light_output_model)
```


### Independence
Based on a plot of standardized residual vs factor level (in this case, glass type and temperature) the residuals not independent of factor level. 

### Normality
Based on a quantile-quantile plot te residuals are very close to normally distrbuted with a mean of zero.

### Homoscedasticity
based on a plot of residual vs fitted values, variation in residual mostly constant except for a cluster of high residuals at $\hat{y}=800$.


# 6
Sludge is the dried product remaining from processed sewage. 
It can be used as fertilizer on crops. 
However, it contains heavy metals. 
They hypothesized the concentration of certain heavy metals in sludge differ among the metropolitan areas from which the sludge is obtained. 
The sludge was aded to the sand at 3 different rates: 0.5,1.0,1.5 metric tons/acre.
The zinc levels were recorded.

## a
Set up a dataframe named metals. Use factos city (A,B,C), rate (0.5,1.0,1.5), and zinc for the observations.
```{r}
cities <- c("A","B","C")
rates <- c("0.5","1.0","1.5")
zinc_data <- expand.grid(rate=rates,city=cities)
zinc_data <- cbind(zinc_data,zinc=c(26.4,25.2,26.0, 30.1,47.7,73.8,
                                    19.4,23.2,18.9, 23.5,39.2,44.6,
                                    31.0,39.1,71.1, 19.3,21.3,19.8,
                                    25.4,25.5,35.5, 30.8,55.3,68.4,
                                    18.7,23.2,19.6, 22.9,31.9,38.6,
                                    32.8,50.7,77.1, 19.0,19.9,21.9
                                    )
                   )
zinc_data %>% kable()
```

## b
Build an aov model, using zinc as the response. 

```{r}
zinc_model <- aov(formula = zinc ~ rate*city ,data=zinc_data)
summary(zinc_model)
```

Which factors are significant? Interaction is significant? 

Rate, City, and the interaction between rate and city are all significant.

Perform an interaction plot.
```{r}
interaction.plot(x.factor = zinc_data$rate,
                 trace.factor = zinc_data$city,
                 response = zinc_data$zinc)
```
Different cities have different response in zinc concentration to increased rate.

## c
List all the factor means and effects. using tapply() or model.table().

```{r}
zinc_model_tables <- model.tables( zinc_model, type = "means" )
zinc_model_tables$tables$"rate:city"
```

## d
calculate the interaction sum squares from scratch.

```{r}
grand_mean <- mean(zinc_data$zinc)
rate_means <- zinc_data %>% group_by(rate) %>% summarise(rate_mean = mean(zinc))
city_means <- zinc_data %>% group_by(city) %>% summarise(city_mean = mean(zinc))
cell_means <- zinc_data %>% group_by(city,rate) %>% summarise(cell_mean = mean(zinc))

zinc_data_means <- zinc_data %>%
  cbind(grand_mean = grand_mean) %>% 
  merge(rate_means,by="rate") %>%
  merge(city_means,by="city") %>%
  merge(cell_means,by=c("rate","city"))

total_df <- nrow(zinc_data) - 1
rate_df <- nrow(rate_means) - 1
city_df <- nrow(city_means) - 1
interaction_df <- rate_df*city_df

total_deviations <- zinc_data_means$zinc - zinc_data_means$grand_mean
total_sum_of_squares <- sum(total_deviations^2)
total_mean_square <- total_sum_of_squares / total_df

rate_deviations <- zinc_data_means$zinc - zinc_data_means$rate_mean
rate_sum_of_squares <- sum(rate_deviations^2)
rate_mean_square <- rate_sum_of_squares / rate_df

city_deviations <- zinc_data_means$zinc - zinc_data_means$city_mean
city_sum_of_squares <- sum(city_deviations^2)
city_mean_square <- city_sum_of_squares / city_df

interaction_deviations <- zinc_data_means$cell_mean - zinc_data_means$city_mean - zinc_data_means$rate_mean + zinc_data_means$grand_mean
interaction_sum_of_squares <- sum(interaction_deviations^2)
interaction_mean_square <- interaction_sum_of_squares/interaction_df
interaction_sum_of_squares
```

the interaction sum of squares is 1809.398 which is consistent with the summary output in part b.