---
title: "Stat 5309 Midterm Project"
author: "Tom Wilson"
date: "Mar 21, 2019"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(agricolae)
library(rsm)
```


# 1
Three brands of batteries are under study. It is suspected that the lives (in weeks) of the three brands are different. Five batteries of each brand are tested with the following results.

```{r}
```
## a
build a linear model, using aov. Are the lives of these brands of batteries different?
which brand gives the longest life?

## b
Perform a boxplot and a stripchart.

## c
construct a 95% CI for the mean life of brand 2. Using Fisher LSD.
Construct a 99% CI for the mean difference between the lives of brand 2 and 3, using Fisher LSD.

## d
perform the 3part residual check.

## e
calculate the number of replicates for a power of 0.9

# 2
A manufacturer of television sets is interested in the effect of tube conductivity of four different types of coating for color picture tubes. The following conductivity data are obtained.

## a
Is there a difference in conductivity due ot coating type?

## b
Estimate the overall mean and the treatment effects.

## c
Compute a 95% CI of the mean of coating type 4.
Compute a 99% CI for the difference between types 1 and 4.

## d
Test all pairs of means using fisher lsd.

# 3
An article in the fire safety journal describes an experiment in which a shape factor was determined for several different nozzle designs at six levels of jet effulx velocity. Interest focused on potential differences between nozzle designs, with velocity considered as a nuisance variable. The data are shown below.

## a
build a linear model nozzle as a blocking factor. Does the nozzle design affect the shape factor?

## b
which nozzle desing are different with respect to the shape factor?

## c
Is the velocity effect significant?

## d
Check if nozzle is an effective blocking factor.

# 4
Johnson and Leone describe an experiment to investigate warping of copper plates. The two factors studied were the temperature and the copper content of the plates. The response variable was a measure of the amount of warping. The data were as follows.

## a
is there any indication that either factor affects the amount of warping?

## b
Do a box plot of warp vs temp and content together.
which level combination gives the lowest warpage?

## c
Suppose that temperature cannot be controlled in where the ocpper plate are used. Any conclusion about the content? (i.e. any difference among means?)
## d
perform a 3-part residual check.


<!--
midterm project
-->

# 1
The effective life of insulating luids at an accelerated load of 35 kV is being studied.
Test data have been obtained for four types of fluids. The results were as follows:

## a
Either read data into R or create the dataframe.

```{r}
fluidtypes <- c("1","2","3","4")
data <- data.frame(fluid = rep(fluidtypes,each=6)
                   ,lifetime=c(17.6,18.9,16.3,17.4,20.1,21.6,
                               16.9,15.3,18.6,17.1,19.5,20.3,
                               21.4,23.6,19.4,18.5,20.5,22.3,
                               19.3,21.1,16.9,17.5,18.3,19.8
                               ))
data %>% kable()
```

## b
Build a linear model, using aov. Is there a significant difference among treatment means?
which fluid gives the longer life?

## c
Construct a 95% Confidence Interval for themean life of fluid 2.
Construct a 99% Confidence Interval for the difference between the lives of Fluids 2 and 3.

## d
Performa complete 3-part residuals check.

## e
Calculate the number of erplicates for a power of 0.99

# 2

## a
Either read data into R or create the dataframe.
```{r}
oils <- c("1","2","3")
trucks <- c("1","2","3","4","5")
data <- expand.grid(truck=trucks,oil=oils)
data <- cbind(data,fuel_consumption = c(0.5,
                                        0.634,
                                        0.487,
                                        0.329,
                                        0.512,
                                        0.535,
                                        0.675,
                                        0.52,
                                        0.435,
                                        0.54,
                                        0.513,
                                        0.595,
                                        0.488,
                                        0.4,
                                        0.51))
data
```

## b
Build a linear model. Is there any significant difference of means about the oil types?
Which oil type gies the lowest fuel consumption?

## c
Is the blocking approach effective?

## d
Do a complete residual assumption check.

# 3
Suppose that in in Problem 4-15, the engineer suspects that the workplaces used by the four operators may represent an additional source of variation. Analyze the data from this experiment (use $\alpha=0.05$) and draw conclusions.

## a
Set up a dataframe with 2 blocking factors (order and operator) and treatment (A,B,C,D)
```{r}
orders <- c("1st","2nd","3rd","4th")
operators <- c("op1","op2","op3","op4")
data <- expand.grid(operator=operators,order_of_assembly=orders)
data <- cbind(data,workplace=c("C","B","D","A",
                               "B","C","A","D",
                               "A","D","B","C",
                               "D","A","C","B"
                               ),
              observation=c(11,10,14,8,
                            8,12,10,12,
                            9,11,7,15,
                            9,8,18,6
                            )
              )
data %>% kable()
```

## b
Use Latin Square to analyze the treatment means.

## c
Which level combination brings the lowest time?

# 4
The factors that influence the breaking strength of a syntheti fiber are being studied. Four production machines and three operators are chosen and a factorial experiment is run using fiber from the same production batch. The results follow.

## a
Either read data into R or create the dataframe.
```{r}
machines <- c("1","2","3","4")
operators <- c("o1","o2","o3")
data <- expand.grid(machine=rep(machines,2),operator=operators)
data <- cbind(data,strength = c(109,110,108,110,
                                110,115,109,108,
                                110,110,111,114,
                                112,111,109,112,
                                116,112,114,120,
                                114,115,119,117
                                )
              )
data %>% kable()
```

## b
Build a linear model. Any interaction between operator and machine?

## c
Build a reduced model.

## d
Do a complete 3-part residual assumption check.

# 5
An experiment is conducted to study the influence of operating temperature and three types of face-plate glass in the light output of an oscilloscope tube. The following data are collected.

## a
Either read data into R or create the dataframe.
```{r}
temperatures <- c(100,125,150)
glasses <- c("t1","t2","t3")
data <- expand.grid(temperature=rep(temperatures,3),
                    glass = glasses)
data <- cbind(data,output=c(580,1090,1392,
                            568,1087,1380,
                            570,1085,1386,
                            
                            550,1070,1328,
                            530,1035,1312,
                            579,1000,1299,
                            
                            546,1045,867,
                            575,1053,904,
                            599,1066,889))
data %>% kable()
```

## b
Build a linear model. Any interaction between glass type and temperature?

## c
Build a reduced model.

## d
Do a complete 3-part residual assumption check.

# 6
Sludge is the dried product remaining from processed sewage. 
It can be used as fertilizer on crops. 
However, it contains heavy metals. 
They hypothesized the concentration of certain heavy metals in sludge differ among the metropolitan areas from which the sludge is obtained. 
The sludge was aded to the sand at 3 different rates: 0.5,1.0,1.5 metric tons/acre.
The zinc levels were recorded.

## a
Set up a dataframe named metals. Use factos city (A,B,C), rate (0.5,1.0,1.5), and zinc for the observations.
```{r}
cities <- c("A","B","C")
rates <- c(0.5,1.0,1.5)
data <- expand.grid(rate=rates,city=cities)
data <- cbind(data,zinc=c(26.4,25.2,26.0, 30.1,47.7,73.8, 19.4,23.2,18.9,
                          23.5,39.2,44.6, 31.0,39.1,71.1, 19.3,21.3,19.8,
                          25.4,25.5,35.5, 30.8,55.3,68.4, 18.7,23.2,19.6,
                          22.9,31.9,38.6, 32.8,50.7,77.1, 19.0,19.9,21.9
                          )
              )
data %>% kable()
```

## b
Build an aov model, using zinc as the response. Which factors are significant? Interaction is significant? Perform an interaction plot.

## c
List all the factor means and effects. using tapply() or model.table().

## d
calculate the interaction sum squares from scratch.












<!-- 
 lab 5 below
-->

# 1.

The yield of a chemical process is being studied. The two most important variables are thought ot be th pressure and the temperature. Three levels of each factor are selected and a factorial experiment with two replicates is performed. The yield data follows.

## a
Set up the dataframe.

```{r}
temperatures <- c("150","160","170")
pressures <- c("200","215","230")
process <- expand.grid(pressure = rep(pressures,2),
                       temperature = temperatures)

process <- cbind(process,yield=c(90.4,90.7,90.2,
                                 90.2,90.6,90.4,
                                 90.1,90.5,89.9,
                                 90.3,90.6,90.1,
                                 90.5,90.8,90.4,
                                 90.7,90.9,90.1))
process %>% kable()
```

## b
Build a linear model using aov(). Are the pressure means significant? Are the temp means significant? Is the interaction significant?

```{r}
yield_model <- aov(yield ~ temperature * pressure, data=process)
summary(yield_model)
```
The interaction between temperature and pressure is not significant.
The main effects of temperature and pressure are significant.

```{r}
yield_model <- aov(yield ~ temperature + pressure, data=process)
summary(yield_model)
```

## c
Create a boxplot of

### yield vs temp
```{r}
boxplot(yield~temperature, data=process)
```

### yield vs pressure
```{r}
boxplot(yield~pressure, data=process)
```

### yield vs temp and pressure
```{r}
boxplot(yield~temperature*pressure, data=process)
```

## d
Perform a residuals assumption check
```{r}
qqnorm(yield_model$residuals)
qqline(yield_model$residuals)
```
The residuals are consistent with a normal distribution centered on zero with constant variance.

# 2
Johnson an dLeone describe an experiment to investigate warping of copper plates. The two factors studied were the temperature and the copper content of the plates. The response variable was a meaure of the amont of warping. The data were as follows.

## a
Set up the dataframe 

```{r}
temperatures <- c("50","75","100","125")
copper_contents <- c("40","60","80","100")
copperplate <- expand.grid(copper_content=rep(copper_contents,2),
                           temperature=temperatures
                           )
copperplate <- cbind(copperplate,warping = c(17,16,24,28,
                                             20,21,22,27,
                                             12,18,17,12,
                                             9,13,12,31,
                                             16,18,25,30,
                                             12,21,23,23,
                                             21,23,23,29,
                                             17,21,22,31))
copperplate %>% kable()
```

## b
build a response model surface (RSM) with warpage as response, use rsm().
```{r}
temp_nums <- copperplate$temperature %>% as.numeric()
cu_nums <- copperplate$copper_content %>% as.numeric()
response <- copperplate$warping
copper_model<- rsm(response ~ SO(temp_nums, cu_nums), data=copperplate)
summary(copper_model)

```
